---
title: "Performance Measure Machine Learning Project"
author: "Eric J Sullivan"
date: "October 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

In this project we have used sensor data from the Human Activity Recognition data set to train and evaluate a random forest model to predict the quality of exercises performed from the measured value of multiple sensor outputs.

<a href = "http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har">http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har</a>

## Data Preparation

First,  load the data.  The pml_testing file becomes the validation data set and the pml_training data set is split to create a training set, pml_trainingshorta, and a testing training set, pml_trainingshortb.  The training and testing sets have both had all variables with NA values removed.

```{r warning=FALSE messages = FALSE}
library(readr)
library(caret)

pml_training <- read_csv("pml-training.csv")
pml_validation <- read_csv("pml-testing.csv")
pml_training$classe <- as.factor(pml_training$classe)

pml_trainingshort <- pml_training[ , colSums(is.na(pml_training))==0 ]

inTrain <- createDataPartition( y = pml_trainingshort$classe, p = 0.7, list = FALSE)
pml_trainingshorta <- pml_trainingshort[inTrain,]
pml_trainingshortb <- pml_trainingshort[-inTrain,]
```

## Train and Evaluate Model

A random forest model is trained on the training data set and its confusion matrix and accuracy for the training set are displayed below.

```{r warning=FALSE}
set.seed(21141141)
PmlRFModel <- train(classe~.,data = pml_trainingshorta[,8:60], method = "rf")
confusionMatrix(PmlRFModel)
```

The model is evaluated against its predictions for the testing data set.

```{r warning=FALSE}
confusionMatrix(pml_trainingshortb$classe,predict(PmlRFModel,pml_trainingshortb))
```

## Predict validation data set

Finally, here are the model's predictions for the validation data set:

```{r warning=FALSE}
PmlRFPrediction <- predict(PmlRFModel,pml_validation)
PmlRFPrediction
```

